Authors Response to Reviewer's Comments

### 2023-10-31 manuscript/review/authorsResponse.qmd
small corrections in the response based on Haralds comments

The objective of running the different t-tests were to allow us for discussing differences between mean values per site between models and observations and between 7 and 10 cm diameter thresholds. The t-test, as employed in our study, is not intended to serve as definitive proof of our findings and we have deliberately avoid over-interpretation of their outcomes. Rather, it serves as a practical tool for managing the extensive number of comparisons in our study. As such, we intend to include a statement in the manuscript to clarify this aspect, please see L\[336-337\].

Following the reviewers comment we included a correction for multiple testing although we are aware that all corrections have pros and cons. In this case we used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller or equal than the nominal test level.

We also look into the alternative approaches suggested by the reviewer. We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case as for example some models did not simulate all the sites. We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale) as suggested. However this approach will not give us the observed-model comparison that we are interested in. Following the reviewers comments we also did further testing to see if with other approaches we could better explain differences between observed and simulated values. We tested a mixed model to account for the randomness induce due to site (plot) differences. For example a glme model with the recruitment values as the response variable and models as the explanatory variable with observed data as the reference (with poisson distribution-count data). This worked well and the patterns were the same as in the t-test (expect for the model 4C). We also considered doing the same approach with the shannon index and mortality differences but in this case we think it is more interesting to understand the mean differences across sites and therefore a t-test is a valid approach. We also tested a linear mixed model to assess differences between 7 and 10cm on the Shannon index creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold, then we run a pairwise multiple comparisons to determine which means differ. The patterns here were also similar to the ones observed in the t-test except for two landscape level models (TreeMig and LandisII) that were significant in this test contrary to the t-test.

After considering all the options that the reviewer suggested and testing others, we made some adjustments to the t-tests analysis, with the correction for multiple testing and more clean output tables as suggested by the reviewer. In addition we explain the purpose and interpretation of these tests better in line \[336-337\].

@@

In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore based on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models. Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data; or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.

@@

The reviewer is correct that we did not provide e.g. site coordinates (with the exception of iLand that needed this for a model specific soil input, this is described in the manuscript), but we did provide topography, climate and soil site specific variables.

@@

Following the reviewers comment we have expanded the caption of table 1.

>>>>

The objective of running the t-tests was to allow us to discuss differences between mean values per site between models, and observations and simulations between the 7 and 10 cm diameter thresholds. The t-test, as employed in our study, is not intended to serve as \'proof\' of our findings, and we have carefully avoided an over-interpretation of their outcomes. Rather, they serve as a practical tool for handling the extensive number of comparisons in our study. To clarify this, we have included a statement in the manuscript, cf. L\[336-337\].

Following the reviewer\'s comment, we have included a correction for multiple tests, although we are aware that all corrections have pros and cons. We used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller than or equal to the nominal test level.

We have also looked into the alternative approaches suggested by the reviewer. We cannot do a paired t-test as for paired samples it is necessary to have the same number of subjects in both groups (observed and simulated), which is not the case here, as for example some models did not simulate all the sites. Yet, we tested the possibility of creating a GLME using as random effect the site and as explanatory variables all the features that we were interested in (model complexity, feedback, model type and scale), as suggested by the reviewer. However, this approach does not provide the comparison of observed vs. simulated data that we are interested in. Following the reviewer\'s comments, we conducted further tests to see if we could better explain differences between observed and simulated values using other approaches. We tested a mixed model to account for the randomness induced by to site (plot) differences. For example, a GLME with recruitment values as the response variable and models as the explanatory variable, with observed data as the reference (with a Poisson distribution because we are dealing with count data). This worked well, and the patterns were the same as in the t-test (except for the model 4C). We also considered pursuing the same approach with the Shannon index and the differences in mortality, but in this case we think it is more relevant to understand the mean differences across sites, and therefore a t-test is more suitable. We furthermore tested a linear mixed model to assess differences between the 7 and 10 cm thresholds regarding the Shannon index, by creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold. We then ran pairwise multiple comparisons to determine which means are different. Also here, the patterns were quite similar to the ones resulting from the t-test except for two landscape-level models (TreeMig and Landis-II) that were significant in this test, contrary to the t-test.

After considering all the options that the reviewer suggested and testing others, we have made some adjustments to the t-test analysis, with the correction for multiple testing and cleaner output tables, as suggested by the reviewer. In addition, we have explained the purpose and interpretation of these tests better on lines \[336-337\].

@@

In this first assessment, we were not aiming to identify what specific subcomponents may hold the key. Therefore, based on the current approach, we could not provide more concrete recommendations for model development, which would most likely be specific for each of the 15 models. Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions. We agree with the reviewer that we need more insights into how we can better simulate regeneration processes in each of these models, but this was beyond the scope of this paper.

@@

The reviewer is correct that we did not provide e.g. site coordinates (with the exception of iLand and aDGVM2 that needed this for a model specific soil input, this is described in the manuscript), but we did provide topography, climate and soil site specific variables.

@@

Following the reviewers comment we have expanded the caption of table 1. We have decided not to include further variables on the model assumptions as we believe the current information provides a good general overview of the models approach and further details are included in the appendix.
